---
title: "Mapping Feature Importance"
author: "Jason McDermott"
date: "7/15/2019"
output: html_document
---

This example will show how to map feature importance (as determined in a simple kmer counting manner)
to individual sequences and plot the results. This can be used to identify regions of the protein that
are most relevant to the function being captured by the positive example set.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load R functions to run SVM cross-validation, etc.
source("SIEVEUb.R")
```

```{r data-ingest}

  # To generate features from a fasta file use the following command
  #    from the SIEVEServer code
  # KmerFeatures.py -f [input.faa] -o [outputbase] -m simple -k 14 -M reduced_alphabet_0
  
  
  # read in data file about examples (class and sequence family)
  ubex_classes = read.table("data/FamiliesConservative.txt", sep="\t", header=1, row.names=1)
  ubex_fact = factor(x=ubex_classes[,3], labels=c("positive", "negative"))
  names(ubex_fact) = rownames(ubex_classes)
  ubex_families = ubex_classes[,1]
  names(ubex_families) = rownames(ubex_classes)
  
    # this is a matrix of features from examples
  ubex_k14red0 = read.table("data/red0/ubligase_k14.red0.train", sep="\t", row.names=1, header=1, stringsAsFactors=F)

  
  # For this example you will also need to generate a list of the kmers
  #   in order for your example protein of interest. Do so using the following
  #   command from the SIEVEServer code
  # KmerFeatures.py -f [input.faa] -o [outputbase] -m simple -k 14 -M reduced_alphabet_0 -K > [output_seqkmer.txt]
  #   
  #    In this case the output file specified in outputbase can be ignored - we'll only use the output_seqkmer.txt
  
  
  seqkmer = read.table("output_seqkmer.txt", sep="\t", row.names=1, stringsAsFactors=F)
  seqkmer[,2] = sapply(seqkmer[,2], function (n) paste("KMER.14.RED0.", substr(n, 2,15), sep=""))
  
```


``` {r matching}

# do the kmercounts on the training set to calculate the importance of each individual kmer
#    in the dataset
kmercounts = family_counts(ubex_k14red0[names(ubex_fact),], ubex_families, ubex_fact)

seqkmercounts = t(sapply(rownames(seqkmer), function (i) {this=seqkmer[i,2]; if (this %in% rownames(kmercounts)) return(c(seqkmer[i,], kmercounts[this,])); return(c(seqkmer[i,], c(NA, NA, NA, NA, NA)))}))

# this gives a data frame with information about each kmer in the sequence - and information about
#   how informative that kmer is to the classification task you have.


# now we can plot kmer information by sequence location
```
Plot of the importance of each position in a sequence based on the contribution of the
kmer at that position to the classification task based on simple counting metrics.

```{r plot-1}
plot(x=rownames(seqkmercounts), y=seqkmercounts[,8], type="l", xlab="Sequence  position", ylab="Importance")

```

Plot of the importance of a window of kmers (smoothed using a window) in a  sequence

```{r plot-2}
plot(sapply(1:(nrow(seqkmercounts)-30), function(i) mean(unlist(seqkmercounts[i:(i+30),8]), na.rm=T)), type="l", xlab="Sequence position", ylab="Mean Importance (window=30)")

```

